---
title: "PML Course Project"
author: "Kamesh Narayanan"
date: "10/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal of the project:
To build a machine learning model to predict the manner in which the participants did their exercises and then predict 20 test cases using the built model.

## Get Data:
Set working directory to the project file

```{r, echo = TRUE}
setwd("C:/Users/MAHE/Desktop/PML")
```
Load caret and randomForest packages

```{r, echo=TRUE}
library(caret)
library(randomForest)
```
Read data into workspace

```{r, echo = TRUE}
rawData <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!", ""))
```

## Clean Data
1. Get rid of columns that have NAs
2. Get rid of columns 1 to 7 since they dont have data useful for analysis
3. Remove data with zero variability
4. Coerce the column 'Classe' to factor

```{r, echo=TRUE}
rawData <- rawData[, colSums(is.na(rawData)) == 0]
rawData <- rawData[, -c(1:7)]
dim(rawData)

noVar <- nearZeroVar(rawData, saveMetrics=TRUE)
data <- rawData[,noVar$nzv==FALSE]
data[, 53] <- as.factor(data$classe)
dim(data)
```

## Split the data into training (75%) and validation (25%) sets 
```{r, echo=TRUE}
inTrain <- createDataPartition(y = data$classe, times = 1, p = 0.75, list = FALSE)
training <- data[inTrain,]
testing <- data[-inTrain,]
```

## Fit a machine learning model and plot it
I have used the function randomForest instead of train because it is faster.
Predcit on the validation set. This is an important set as we can find the out of sample error.
Print the confusion matrix to show error

```{r, echo=TRUE}
RFmodel <- randomForest(classe ~ ., data = training)
RFprediction <- predict(RFmodel, testing) #Cross validation
RFconfMat <- confusionMatrix(RFprediction, testing$classe)
RFconfMat
plot(RFmodel)
```

## Predict on the given test set
```{r, echo = TRUE}
TestingSET <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!", ""))
finalPrediction <- predict(RFmodel, TestingSET)
finalPrediction
```